Parallel Computing for Constrained Global Optimization Problems \thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

### 2018-05-04 2018/Journal of Supercomputing/JoS_Paper.tex
Before proofreading

\begin{abstract}
This work considers a parallel algorithm for solving multiextremal problems with non-convex constraints. The distinctive feature of this algorithm, which does not use penalty functions, is the separate consideration of each problem constraint. The search process can be conducted by reducing the original multidimensional problem to a number of related one-dimensional problems and solving this set of problems in parallel. An experimental assessment of parallel algorithm efficiency was conducted by finding the numeric solution to several hundred randomly generated multidimensional multiextremal problems with non-convex constraints.
\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

This work considers parallel methods for solving global optimization problems with non-convex constraints. The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the problem’s functions may also be unknown, i.e. they can be set by an algorithm computing their values at various points within the search domain (so-called ``black-box'' functions). Moreover, it is supposed that even a single computation of the function value can be time-consuming, as in the applied problems it requires performing numerical simulation. These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.

At the same time, it is common for applied constrained optimization problems to be in a situation where violating one constraint results in all other functions returning indeterminate values. An example includes optimal control problems, described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}. It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e. every eigenvalue of $A$ has strictly negative real part. Otherwise the value of the criteria is indeterminate.

This partial computability of the functions in constrained optimization problems substantially complicates the application of the well-known penalty function method (in some cases making it completely impossible). Thanks to its simplicity, this method is one of the most popular approaches to solving problems with constraints. However, calculating the penalty function requires first calculating the values of all of the problem’s functions at the given point, which is impossible if they are partially indeterminate.

In this work, the authors consider an approach to minimizing multiextremal functions under non-convex constraints, developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002} and called \textit{the index method}. The approach is based on separate considering every constraint in the problem and is not related to using penalty functions. According to the index method, each iteration (\textit{a trial}) at a respective point in the search domain includes a sequential check of the problem constraints at that point. As soon as the first constraint violation is found, the trial is interrupted and the method proceeds with the next iteration; no other problem functions are calculated at that point. This allows problems to be solved in which function values may not be determined for the entire search domain. Under this approach, solving multidimensional problems is reduced (using Peano-type space-filling curves) to solving equivalent one-dimensional problems.

It should be noted that standard approaches to algorithm parallelization are not quite applicable to global optimization. For example, the rules for selecting another iteration point are quite simple and do not require parallelization (as overheads associated with organizing parallel computations will nullify any possible acceleration). Some acceleration can be achieved by parallelizing the computation of function values describing the object to be optimized; however, this approach is specific to each individual problem being solved.

The following approach looks more promising. The algorithm can be modified to run several trials in parallel. This approach provides the efficiency (as parallelization is applied to the most computation-intensive part of the problem solving process) and generality (in that it applies to a wide range of global optimization algorithms). The approach, described in \cite{Sidorov2015} for unconstrained optimization, was used in this work for parallelizing constrained optimization algorithms.

@@

Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
\end{equation}
The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy Lipschitz condition 
\[
\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
\]
with a priory unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is admitted that functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
\begin{equation}\label{g_k}
g_k(y) \leq 0, \; 1 \leq k < i.
\end{equation}

@@

>>>>

\begin{abstract}
In the present work, a parallel algorithm for solving the multiextremal problems with non-convex constrains is considered. Within the framework of approach considered, the solving of initial multidimensional problem is reduced to the parallel solving of a set of linked one-dimensional problems. The parallelization scheme is featured by the absence of a single control process since the solution of any one-dimensional problem is a solution of the initial multidimensional one. The non-convex constraints are accounted for by a novel index scheme not utilizing the ideas of penalty functions. The efficiency of the parallel algorithm was confirmed experimentally by the numerical solving of several hundred multidimensional multiextremal problems with non-convex constrains generated randomly.
\keywords{Global optimization \and Constrained problems \and Non-convex constraints \and Dimensionality reduction \and Parallel algorithms}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

This work considers parallel methods for solving global optimization problems with non-convex constraints. The objective function and constraints are assumed to satisfy the Lipschitz condition with a priori unknown Lipschitz constants. The analytical form of the problem’s functions may also be unknown, i.e. they can be set by an algorithm computing their values at various points within the search domain (so-called ``black-box'' functions). Moreover, it is supposed that even a single computation of the function value can be time-consuming, as in the applied problems it requires performing numerical simulation. 
%These assumptions are typical for many approaches to building parallel algorithms for unconstrained global optimization \cite{Jones2001,Paulavicius2011,Evtushenko2013}.

The numerical algorithms for solving the problems of the specified class are featured by an exponential growth of the computational costs with increasing dimensionality (so called ``curse of dimensionality''). The use of supercomputers expands the field of application of the global optimization methods and, at the same time, sets a problem of efficient parallelization of the search process.

At present, the parallel algorithms, which utilize the heuristic approaches and based on the idea of random search in any way are quite popular (see \cite{RefFerreiro,RefZhu,Korosec,Guerrero}). However, because of the probabilistic nature, the algorithms of such type guarantee the convergence to the global optimum solution in the probabilistic meaning only, that distinguishes these ones unfavorably from the deterministic parallel global optimization algorithms \cite{Jones2001,Paulavicius2011,Evtushenko2013}. The efficient deterministic algorithms are oriented onto the solving of the unconstrained optimization problems as a rule. The problems with constraints are solved using the penalty function method usually having a number of disadvantages. One of these is the impossibility of its application in the problems with the \textit{partially defined functions} (i.e. when violating any constraint, the values of the rest functions of the problem remain undefined).

As the examples, one can mention the optimal control problems, described through systems of ordinary differential equations with a certain matrix $A$ on the right side \cite{Balandin2011,Balandin2017}. It is only possible to calculate the optimality criteria for these problems if the matrix $A$ is a Hurwitz matrix, i.e. every eigenvalue of $A$ has strictly negative real part. Otherwise the value of the criteria is indeterminate.

In the present work, a novel approach to the minimization of the multiextremal functions with non-convex constraints developed in \cite{Strongin2000,Sergeyev2001,Barkalov2002} and referred to as the index method  of accounting for the constraints was applied. The approach is based on a separate accounting for each constraint of the problem and is not related to the use of the penalty functions. The solving of initial multidimensional problem is reduced (using the Peano-type space-filling curves) to solving an equivalent set of the one-dimensional problems, which can be solved in parallel in such a way that each of these processes shares the information obtained by the other processes. The propose scheme was implemented using Lobachevsky supercomputer with the use of the distributed memory as well as the shared one.

@@

Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y):y\in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}},
\end{equation}
\begin{equation}\label{D}
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
\end{equation}
The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_i(y), \; 1\leq i \leq m,$ of the constraints satisfy Lipschitz condition 
\[
\left|g_i(y_1)-g_i(y_2)\right|\leq L_i\left\|y_1-y_2\right\|, \;1\leq i\leq m+1, \; y_1,y_2 \in D,\;
\]
with a priory unknown constants $L_i, \; 1 \leq i \leq m+1,$ and may be multiextremal. It is assumed that functions $g_i(y)$ are defined and computable only at the points $y \in D$ satisfying the conditions
\begin{equation}\label{g_k}
g_k(y) \leq 0, \; 1 \leq k < i.
\end{equation}

@@

Computational experiments were carried out on a high-performance cluster of Lobachevsky State University of Nizhni Novgorod. The cluster node included two Intel Sandy Bridge E5-2660 2.2 GHz CPUs and 64 Gb RAM. The CPU had 8 cores, i.e. 16 cores were available in each node.
